--
title: "Reproducible Research - Assessment 1"
author: "José Miguel Reséndiz"
date: "Tuesday, February 10, 2015"
output: html_document
---

##INTRODUCCTION

  It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a [Fitbit](http://www.fitbit.com/mx), [Nike Fuelband](http://www.nike.com/us/en_us/c/nikeplus-fuelband), or [Jawbone Up](https://jawbone.com/intl/up). These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks.
  But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

##Data
The data for this assignment can be downloaded from the course web site:


Dataset: [Activity monitoring data [52K]](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)


The variables included in this dataset are:


**steps:** Number of steps taking in a 5-minute interval (missing values are coded as NA)


**date:** The date on which the measurement was taken in YYYY-MM-DD format


**interval:** Identifier for the 5-minute interval in which measurement was taken


The dataset is stored in a comma-separated-value (CSV) file and there are a
total of 17,568 observations in this dataset.

###LOADING AND PROCESSING THE DATA

Load Data
```{r simulation}
dataPerMov<-read.csv("activity.csv")
```
Raw Data
```{r,echo=FALSE}
head(dataPerMov)
```
Structure
```{r,echo=FALSE}
str(dataPerMov)
```

Select only the rows with values different of NA.
```{r simulatio}
library(stringr)
dataPerMov<-read.csv("activity.csv")
dataPerMov<-na.omit(dataPerMov)
dataPerMov <- data.frame(date = dataPerMov$date,  interval = dataPerMov$interval, steps = dataPerMov$steps)
head(dataPerMov)
```

###What is mean total number of steps taken per day?

Histogram of the total number of steps taken each day
``` {r,echo=FALSE,fig.width=8, fig.height=8 }
library(stringr)
library(ggplot2)
dataPerMov<-read.csv("activity.csv")
dataPerMov<-na.omit(dataPerMov)
dataPerMov <- data.frame(date = dataPerMov$date, interval = dataPerMov$interval, steps = dataPerMov$steps)
stepsForDay<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=sum)
ggplot(stepsForDay, aes(x=date, y=x)) + geom_bar(stat="identity") + 
  labs(x="Dates", y="Stpes") + 
  coord_flip()
```

Mean and median total number of steps taken per day
```{r,echo=FALSE}
library(stringr)
dataPerMov<-read.csv("activity.csv")
dataPerMov<-na.omit(dataPerMov)
dataPerMov <- data.frame(date = dataPerMov$date,interval = dataPerMov$interval, steps = dataPerMov$steps)

stepsForDayMean<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=mean)
stepsForDayMedian<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=median)$x
dataPerMov <- data.frame(DATE = stepsForDayMean$date,MEAN_STEP = stepsForDayMean$x,MEDIAN_STEP=stepsForDayMedian)
dataPerMov
```

###What is the average daily activity pattern?
Series of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
``` {r,echo=FALSE,fig.width=11, fig.height=4 }
library(ggplot2)
library(stringr)
dataPerMov<-read.csv("activity.csv")
dataPerMov<-na.omit(dataPerMov)
dataPerMov <- data.frame(date = dataPerMov$date, interval = dataPerMov$interval, steps = dataPerMov$steps)
stepsForInterval<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$interval), FUN=mean)
interval <- substr(as.character(seq(as.POSIXct("2015-02-12 00:00:00 CST"), as.POSIXct("2015-02-12 23:55:00 CST"), by = "hours")),12,16)
names(stepsForInterval) <- c("INTERVAL","AVERAGE_STEPS")
longuitud <- nrow(stepsForInterval) -1
longuitudInterval <- length(interval)
ggplot(stepsForInterval, aes(0:longuitud, AVERAGE_STEPS)) + geom_line() +
 scale_x_continuous(name="time" ,  breaks=seq(0,longuitud,length.out=longuitudInterval), labels=interval)
```

Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r,echo=FALSE}
dataPerMov<-read.csv("activity.csv")
dataPerMov<-na.omit(dataPerMov)
numRow<-nrow(dataPerMov)
stepsForDayMean<- aggregate(dataPerMov$steps, by=list(interval=dataPerMov$interval), FUN=mean)
dataPerMov <- data.frame(INTERVAL = stepsForDayMean$interval,MEAN_STEP = stepsForDayMean$x)

stepsForInterval<-dataPerMov[max(dataPerMov$MEAN_STEP) == dataPerMov$MEAN_STEP,]
stepsForInterval
```


###IMPUTING MISSING VALUES


Number of NA rows  
```{r,echo=TRUE}
dataPerMov<-read.csv("activity.csv")
nrow(dataPerMov[is.na(dataPerMov$steps),])
```

Filling in all of the missing values in the dataset.We use the mean for that day and 5-minute interval
```{r,echo=TRUE}
#Prepare Data
dataPerMovRaw<-read.csv("activity.csv")
auxData <- cbind(dataPerMovRaw,wday = as.POSIXlt(dataPerMovRaw$date)$wday)
dataPerMov<-auxData[!is.na(auxData$steps),]
dataPerMovNA <-auxData[is.na(auxData$steps),]

#Calculating of mean by interval and week day for the values NA
dataPerMovMean<- aggregate(dataPerMov$steps, by=list(interval=dataPerMov$interval, wday=dataPerMov$wday), FUN=mean)
names(dataPerMovMean) <- c("interval","wday","mean")

#Building final data set
tempAux <- merge(dataPerMovNA,dataPerMovMean)
dataPerMovNA <- data.frame(steps = tempAux$mean, date = tempAux$date, interval = tempAux$interval, wday = tempAux$wday)
finalDataSet <- rbind(dataPerMovNA,dataPerMov)
head(finalDataSet)
```

Output
``` {r,echo=FALSE,fig.width=11, fig.height=4 }
library(ggplot2)
library(stringr)

#Load data
dataPerMovRaw<-read.csv("activity.csv")
auxData <- cbind(dataPerMovRaw,wday = as.POSIXlt(dataPerMovRaw$date)$wday)
dataPerMov<-auxData[!is.na(auxData$steps),]
dataPerMovNA <-auxData[is.na(auxData$steps),]

#Calculation mean by interval and week day for the values NA
dataPerMovMean<- aggregate(dataPerMov$steps, by=list(interval=dataPerMov$interval, wday=dataPerMov$wday), FUN=mean)
names(dataPerMovMean) <- c("interval","wday","mean")

#Building final data set
tempAux <- merge(dataPerMovNA,dataPerMovMean)
dataPerMovNA <- data.frame(steps = tempAux$mean, date = tempAux$date, interval = tempAux$interval, wday = tempAux$wday)
finalDataSet <- rbind(dataPerMovNA,dataPerMov)

#Make Plot
dataPerMov<-finalDataSet
dataPerMov <- data.frame(date = dataPerMov$date, interval = dataPerMov$interval, steps = dataPerMov$steps)
stepsForInterval<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$interval), FUN=mean)
interval <- substr(as.character(seq(as.POSIXct("2015-02-12 00:00:00 CST"), as.POSIXct("2015-02-12 23:55:00 CST"), by = "hours")),12,16)
names(stepsForInterval) <- c("INTERVAL","AVERAGE_STEPS")
longuitud <- nrow(stepsForInterval) -1
longuitudInterval <- length(interval)
ggplot(stepsForInterval, aes(0:longuitud, AVERAGE_STEPS)) + geom_line() +
  scale_x_continuous(name="time" ,  breaks=seq(0,longuitud,length.out=longuitudInterval), labels=interval)
```

###Mean and median total number of steps taken per day
```{r,echo=FALSE}
#Load data
dataPerMovRaw<-read.csv("activity.csv")
auxData <- cbind(dataPerMovRaw,wday = as.POSIXlt(dataPerMovRaw$date)$wday)
dataPerMov<-auxData[!is.na(auxData$steps),]
dataPerMovNA <-auxData[is.na(auxData$steps),]

#Calculation mean by interval and week day for the values NA
dataPerMovMean<- aggregate(dataPerMov$steps, by=list(interval=dataPerMov$interval, wday=dataPerMov$wday), FUN=mean)
names(dataPerMovMean) <- c("interval","wday","mean")

#Building final data set
tempAux <- merge(dataPerMovNA,dataPerMovMean)
dataPerMovNA <- data.frame(steps = tempAux$mean, date = tempAux$date, interval = tempAux$interval, wday = tempAux$wday)
finalDataSet <- rbind(dataPerMovNA,dataPerMov)

dataPerMov <-finalDataSet

stepsForDayMean<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=mean)
stepsForDayMedian<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=median)$x
dataPerMov <- data.frame(DATE = stepsForDayMean$date,MEAN_STEP = stepsForDayMean$x,MEDIAN_STEP=stepsForDayMedian)
dataPerMov
```


###Do these values differ from the estimates from the first part of the assignment?
```{r,echo=FALSE}
#####################################################################
#######CALCULATE VALUES NA
#####################################################################

#Load data
dataPerMovRaw<-read.csv("activity.csv")
auxData <- cbind(dataPerMovRaw,wday = as.POSIXlt(dataPerMovRaw$date)$wday)
dataPerMov<-auxData[!is.na(auxData$steps),]
dataPerMovNA <-auxData[is.na(auxData$steps),]

#Calculation mean by interval and week day for the values NA
dataPerMovMean<- aggregate(dataPerMov$steps, by=list(interval=dataPerMov$interval, wday=dataPerMov$wday), FUN=mean)
names(dataPerMovMean) <- c("interval","wday","mean")

#Building final data set
tempAux <- merge(dataPerMovNA,dataPerMovMean)
dataPerMovNA <- data.frame(steps = tempAux$mean, date = tempAux$date, interval = tempAux$interval, wday = tempAux$wday)
finalDataSet <- rbind(dataPerMovNA,dataPerMov)
dataPerMov <-finalDataSet
stepsForDayMean<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=mean)
stepsForDayMedian<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=median)$x
dataPerMovNA <- data.frame(DATE = stepsForDayMean$date,MEAN_STEP_NA = stepsForDayMean$x,MEDIAN_STEP_NA=stepsForDayMedian)

#####################################################################
#######CALCULATE
#####################################################################
library(stringr)
dataPerMov<-read.csv("activity.csv")
dataPerMov<-na.omit(dataPerMov)
dataPerMov <- data.frame(date = dataPerMov$date,interval = dataPerMov$interval, steps = dataPerMov$steps)

stepsForDayMean<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=mean)
stepsForDayMedian<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=median)$x
dataPerMov <- data.frame(DATE = stepsForDayMean$date,MEAN_STEP = stepsForDayMean$x,MEDIAN_STEP=stepsForDayMedian)

merge(dataPerMovNA,dataPerMov)


```

####What is the impact of imputing missing data on the estimates of the total daily number of steps?
There´re some rows with any value, but this reason, we cant't calculat a value.
```{r,echo=FALSE}
#####################################################################
#######CALCULATE VALUES NA
#####################################################################

#Load data
dataPerMovRaw<-read.csv("activity.csv")
auxData <- cbind(dataPerMovRaw,wday = as.POSIXlt(dataPerMovRaw$date)$wday)
dataPerMov<-auxData[!is.na(auxData$steps),]
dataPerMovNA <-auxData[is.na(auxData$steps),]

#Calculation mean by interval and week day for the values NA
dataPerMovMean<- aggregate(dataPerMov$steps, by=list(interval=dataPerMov$interval, wday=dataPerMov$wday), FUN=mean)
names(dataPerMovMean) <- c("interval","wday","mean")

#Building final data set
tempAux <- merge(dataPerMovNA,dataPerMovMean)
dataPerMovNA <- data.frame(steps = tempAux$mean, date = tempAux$date, interval = tempAux$interval, wday = tempAux$wday)
finalDataSet <- rbind(dataPerMovNA,dataPerMov)
dataPerMov <-finalDataSet
stepsForDayMean<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=mean)
stepsForDayMedian<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=median)$x
dataPerMovNA <- data.frame(DATE = stepsForDayMean$date,MEAN_STEP_NA = stepsForDayMean$x,MEDIAN_STEP_NA=stepsForDayMedian)

#####################################################################
#######CALCULATE
#####################################################################
library(stringr)
dataPerMov<-read.csv("activity.csv")
dataPerMov<-na.omit(dataPerMov)
dataPerMov <- data.frame(date = dataPerMov$date,interval = dataPerMov$interval, steps = dataPerMov$steps)

stepsForDayMean<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=mean)
stepsForDayMedian<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=median)$x
dataPerMov <- data.frame(DATE = stepsForDayMean$date,MEAN_STEP = stepsForDayMean$x,MEDIAN_STEP=stepsForDayMedian)

res <- merge(dataPerMovNA,dataPerMov,all.x=TRUE, all.y = TRUE)
res<-res[is.na(res$MEDIAN_STEP),]
res
```

####Are there differences in activity patterns between weekdays and weekends?

There´re some rows with any value, but this reason, we cant't calculat a value.

```{r,echo=FALSE}
#####################################################################
#######CALCULATE VALUES NA
#####################################################################

#Load data
dataPerMovRaw<-read.csv("activity.csv")
auxData <- cbind(dataPerMovRaw,wday = as.POSIXlt(dataPerMovRaw$date)$wday)
dataPerMov<-auxData[!is.na(auxData$steps),]
dataPerMovNA <-auxData[is.na(auxData$steps),]

#Calculation mean by interval and week day for the values NA
dataPerMovMean<- aggregate(dataPerMov$steps, by=list(interval=dataPerMov$interval, wday=dataPerMov$wday), FUN=mean)
names(dataPerMovMean) <- c("interval","wday","mean")

#Building final data set
tempAux <- merge(dataPerMovNA,dataPerMovMean)
dataPerMovNA <- data.frame(steps = tempAux$mean, date = tempAux$date, interval = tempAux$interval, wday = tempAux$wday)
finalDataSet <- rbind(dataPerMovNA,dataPerMov)
dataPerMov <-finalDataSet
stepsForDayMean<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=mean)
stepsForDayMedian<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=median)$x
dataPerMovNA <- data.frame(DATE = stepsForDayMean$date,MEAN_STEP = stepsForDayMean$x,MEDIAN_STEP=stepsForDayMedian)

```


####Are there differences in activity patterns between weekdays and weekends?
Create a new factor variable in the dataset with two levels - "weekday" and "weekend" indicating whether a given date is a weekday or weekend day
```{r,echo=FALSE}
#####################################################################
#######CALCULATE VALUES NA
#####################################################################

#Load data
dataPerMovRaw<-read.csv("activity.csv")
auxData <- cbind(dataPerMovRaw,wday = as.POSIXlt(dataPerMovRaw$date)$wday)
dataPerMov<-auxData[!is.na(auxData$steps),]
dataPerMovNA <-auxData[is.na(auxData$steps),]

#Calculation mean by interval and week day for the values NA
dataPerMovMean<- aggregate(dataPerMov$steps, by=list(interval=dataPerMov$interval, wday=dataPerMov$wday), FUN=mean)
names(dataPerMovMean) <- c("interval","wday","mean")

#Building final data set
tempAux <- merge(dataPerMovNA,dataPerMovMean)
dataPerMovNA <- data.frame(steps = tempAux$mean, date = tempAux$date, interval = tempAux$interval, wday = tempAux$wday)
finalDataSet <- rbind(dataPerMovNA,dataPerMov)

auxFlag <- finalDataSet$wday==6 | finalDataSet$wday==0
finalDataSet$wday[auxFlag] <- "weekend"
finalDataSet$wday[!auxFlag] <- "weekday"
head(finalDataSet,n=10)
```

Make a panel plot containing a time series plot of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis)

``` {r,echo=FALSE,fig.width=11, fig.height=4 }
#####################################################################
#######CALCULATE VALUES NA
#####################################################################

#Load data
dataPerMovRaw<-read.csv("activity.csv")
auxData <- cbind(dataPerMovRaw,wday = as.POSIXlt(dataPerMovRaw$date)$wday)
dataPerMov<-auxData[!is.na(auxData$steps),]
dataPerMovNA <-auxData[is.na(auxData$steps),]

#Calculation mean by interval and week day for the values NA
dataPerMovMean<- aggregate(dataPerMov$steps, by=list(interval=dataPerMov$interval, wday=dataPerMov$wday), FUN=mean)
names(dataPerMovMean) <- c("interval","wday","mean")

#Building final data set
tempAux <- merge(dataPerMovNA,dataPerMovMean)
dataPerMovNA <- data.frame(steps = tempAux$mean, date = tempAux$date, interval = tempAux$interval, wday = tempAux$wday)
finalDataSet <- rbind(dataPerMovNA,dataPerMov)

auxFlag <- finalDataSet$wday==6 | finalDataSet$wday==0
finalDataSet$wday[auxFlag] <- "weekend"
finalDataSet$wday[!auxFlag] <- "weekday"


finalDataSet <-aggregate(finalDataSet$steps, by=list(interval=finalDataSet$interval, wday=finalDataSet$wday), FUN=mean)
names(finalDataSet) <- c("interval"  ,"wday" , "mean")

interval <- substr(as.character(seq(as.POSIXct("2015-02-12 00:00:00 CST"), as.POSIXct("2015-02-12 23:55:00 CST"), by = "hours")),12,16)
longuitud <- nrow(finalDataSet) -1
longuitudInterval <- length(interval)


ggplot(finalDataSet,  aes(x=interval,  y=mean,  colour=wday,  group=wday)) +
  geom_line() 
```
