---
title: "Reproducible Research - Assessment 1"
author: "José Miguel Reséndiz"
date: "Tuesday, February 10, 2015"
output: pdf_document
---

##INTRODUCCTION

It is now possible to collect a large amount of data about personal movement using activity monitoring devices such as a [Fitbit](http://www.fitbit.com/mx), [Nike Fuelband](http://www.nike.com/us/en_us/c/nikeplus-fuelband), or [Jawbone Up](https://jawbone.com/intl/up). These type of devices are part of the "quantified self" movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks.
But these data remain under-utilized both because the raw data are hard to obtain and there is a lack of statistical methods and software for processing and interpreting the data.

##Data
The data for this assignment can be downloaded from the course web site:


Dataset: [Activity monitoring data [52K]](https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip)


The variables included in this dataset are:


**steps:** Number of steps taking in a 5-minute interval (missing values are
coded as NA)


**date:** The date on which the measurement was taken in YYYY-MM-DD
format


·**interval:** Identifier for the 5-minute interval in which measurement was
taken


The dataset is stored in a comma-separated-value (CSV) file and there are a
total of 17,568 observations in this dataset.

## GETING & CLEANING DATA

###1.- LOAD DATA
Load Data
```{r simulation}
dataPerMov<-read.csv("activity.csv")
```
Raw Data
```{r,echo=FALSE}
head(dataPerMov)
```
Structure
```{r,echo=FALSE}
str(dataPerMov)
```


###2.- Process/transform the data
We select only the rows with values different of NA, and tranform field 'date' to POSIXlt
```{r simulatio}
dataPerMov<-na.omit(dataPerMov)
dataPerMov$date <- as.POSIXlt(dataPerMov$date)
```
Output
```{r,echo=FALSE}
head(dataPerMov)
```

##EXPLORATORY DATA ANALYSI

###1.- What is mean total number of steps taken per day?

``` {r,echo=FALSE,fig.width=8, fig.height=8 }
dataPerMov<-read.csv("activity.csv")
dataPerMov<-na.omit(dataPerMov)
stepsForDay<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=sum)

library(ggplot2)
ggplot(stepsForDay, aes(x=date, y=x)) + geom_bar(stat="identity") + 
  labs(x="Dates", y="Stpes") + 
  coord_flip()
```


``` {r,echo=FALSE,results='asis',fig.align='center'}
dataPerMov<-read.csv("activity.csv")
dataPerMov<-na.omit(dataPerMov)
stepsForDay<- aggregate(dataPerMov$steps, by=list(date=dataPerMov$date), FUN=mean)
library("xtable")
stepsForDay<- xtable(stepsForDay)

print(stepsForDay,type="html")
```

